---
title: "Day 9: Rope Bridge"
author: "@Andi@tech.lgbt"
output: 
  html_notebook: 
    code_folding: none
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
dat <- read.table("aoc09.txt", sep = " ") |>
  rename(dir = V1, n = V2)
dat |> head()
```


I reckon it will be handier if each move is one step, so let's expand the move list...

```{r}
expand_moves <- function(dir, n) {
  rep(dir, n)
}
```

(That didn't really need a function, did it? I'd forgotten that `rep` takes vectors. Onwards.)

```{r}
moves <- tibble(d = expand_moves(dat$dir, dat$n))
moves |> head()
```


It will also be handy to tell if H and T are touching. Voila:

```{r}
touching <- function(p1_x, p1_y, p2_x, p2_y) {
  p1 <- c(p1_x, p1_y)
  p2 <- c(p2_x, p2_y) 
  dist(rbind(p1, p2),
       method = "euclidean") < 2
}
```

(Why 2? It's probably closer to $\sqrt{2}$, but 2 works fine.)

Examples where they touch:

```{r}
touching( 0, 0,  0,0)
touching( 1, 0,  0,0)
touching(-1, 0,  0,0)
touching( 0, 1,  0,0)
touching( 0,-1,  0,0)
touching( 1, 1,  0,0)
touching(-1, 1,  0,0)
touching( 1, 1,  0,0)
touching( 1,-1,  0,0)
```

Examples where they don't:

```{r}
touching( 2, 0,  0,0)
touching(-2, 0,  0,0)
touching( 0, 2,  0,0)
touching( 0,-2,  0,0)
```


This is surely hidden in R somewhere already:

```{r}
bound <- function(x, abs_max) {
  bound_up   <- ifelse(x > abs_max, 
                       abs_max,
                       x)
  bound_down <- ifelse(bound_up < -abs_max,
                       -abs_max,
                       bound_up)
  bound_down
}
```

```{r}
bound(3, 2)
bound(-3, 2)
```


The main function. Note the secret ingredient commented therein.

```{r}
move <- function(start_x, start_y, ms) {
  res <- ms |>
    mutate(h_x = NA,
           h_y = NA,
           t_x = NA,
           t_y = NA)
  
  h_x <- start_x
  h_y <- start_y
  t_x <- start_x
  t_y <- start_y
  
  for (i in seq_along(res$d)) {
    # Move H
    if (res$d[i] == "U")
      h_y <- h_y + 1
    else if (res$d[i] == "D")
      h_y <- h_y - 1
    else if (res$d[i] == "L")
      h_x <- h_x - 1
    else if (res$d[i] == "R")
      h_x <- h_x + 1
    else
      stop("Don't know that direction")
    
    # secret ingredient
    if (!touching(h_x, h_y, t_x, t_y)) {
      delta_x <- (h_x - t_x) |> bound(1)
      delta_y <- (h_y - t_y) |> bound(1)
      t_x <- t_x + delta_x
      t_y <- t_y + delta_y
    }
    
    # Update table
    res$h_x[i] <- h_x
    res$h_y[i] <- h_y
    res$t_x[i] <- t_x
    res$t_y[i] <- t_y
  }
  
  res
}
```


```{r}
moved <- move(0, 0, moves)
```


```{r}
moved |> head()
```

## Part 1:

How many places has T been?

```{r}
paste(moved$t_x, moved$t_y) |>
  unique() |>
  length()
```

Correct!

## Part 2:

Okay, now I wish I'd guessed this was coming. I'm just going to hike out the secret ingredient and do it *n* times, treating the previous knot's moves as a head.

```{r}
follow_n <- function(m, n) {
  res <- m
  stopifnot(n >= 1)
  
  for (k in 1:n) {
    h <- res |> select(ncol(res) - 1, ncol(res))
    names(h) <- c("x", "y")
    h$t_x <- NA
    h$t_y <- NA
    
    t_x <- 0
    t_y <- 0
    
    for (i in 1:nrow(h)) {
      h_x <- h$x[i]
      h_y <- h$y[i]
      
      if (!touching(h_x, h_y, t_x, t_y)) {
        delta_x <- (h_x - t_x) |> bound(1)
        delta_y <- (h_y - t_y) |> bound(1)
        t_x <- t_x + delta_x
        t_y <- t_y + delta_y
      }

      h$t_x[i] <- t_x
      h$t_y[i] <- t_y
    }
    
    prefix <- paste0("t",k,"_")
    names(h)[3:4] <- c(paste0(prefix,"x"),
                       paste0(prefix,"y"))
    
    res <- cbind(res, h[,3:4])
  }
  
  res
}
```

```{r}
moved2 <- moved |> select(d, h_x, h_y)
moved2 <- follow_n(moved2, 9)
```


```{r}
names(moved2)
```


How many places has the 9th knot been?

```{r}
paste(moved2$t9_x, moved2$t9_y) |>
  unique() |>
  length()
```

It worked!
