---
title: "Day 2: Rock Paper Scissors"
author: "@Andi@tech.lgbt"
output: 
  html_notebook: 
    code_folding: none
---

```{r}
library(tidyverse)
```

```{r}
raw_dat <- read.csv("aoc02.txt",
                    blank.lines.skip = FALSE,
                    header = FALSE,
                    sep = " ") |>
  rename(them = V1,
         me = V2)

raw_dat
```

Them: A for Rock, B for Paper, and C for Scissors

Me: X for Rock, Y for Paper, and Z for Scissors


```{r}
move_names <- c("rock", "paper", "scissors")
```


"The score for a single round is the score for the shape you selected (1 for Rock, 2 for Paper, and 3 for Scissors) plus the score for the outcome of the round (0 if you lost, 3 if the round was a draw, and 6 if you won)."


```{r}
scorer <- expand.grid(them_name = move_names,
            me_name = move_names) |>
  mutate(
    outcome_me = c(
      "draw",
      "lose",
      "win",
      "win",
      "draw",
      "lose",
      "lose",
      "win",
      "draw"
    ),
    them = rep(LETTERS[1:3], 3),
    me   = c(rep("X", 3),
             rep("Y", 3),
             rep("Z", 3)),
    outcome_score_me = case_when(
      outcome_me == "lose" ~ 0,
      outcome_me == "draw" ~ 3,
      outcome_me == "win"  ~ 6),
    move_score_me = case_when(
      me_name == "rock"     ~ 1,
      me_name == "paper"    ~ 2,
      me_name == "scissors" ~ 3),
    total_score_me = outcome_score_me + move_score_me
  )
```



```{r}
dat <- left_join(raw_dat, scorer)
dat
```


Part 1:

```{r}
sum(dat$total_score_me)
```

Part 2:

"Anyway, the second column says how the round needs to end: X means you need to lose, Y means you need to end the round in a draw, and Z means you need to win. Good luck!"


```{r}
raw_dat_2 <- raw_dat |>
  mutate(outcome_me = case_when(me == "X" ~ "lose",
                                me == "Y" ~ "draw",
                                me == "Z" ~ "win")) |>
  select(-me)
raw_dat_2
```

```{r}
dat2 <- left_join(raw_dat_2, scorer)
dat2 
```

```{r}
sum(dat2$total_score_me)
```
