---
title: "Day 6: Tuning Trouble"
author: "@Andi@tech.lgbt"
output: 
  html_notebook: 
    code_folding: none
---


```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
dat <- read_lines("aoc06.txt")
```

```{r}
str2vec <- function(s) {
  strsplit(s,"")[[1]]
}
```

```{r}
vec <- str2vec(dat)
```

```{r}
all_unique <- function(x) {
  (x |> table() |> max()) == 1
}
```

```{r}
first_unique <- function(x, seqlen) {
  x_n   <- length(x)
  found <- FALSE
  start <- 1
  end   <- start + seqlen - 1
  
  while(end <= x_n && !found) {
    subseq <- x[start:end]
    if (all_unique(subseq)) {
      found <- TRUE
    }
    else {
      start <- start + 1
      end   <- start + seqlen - 1
    }
  }
  
  if (found)
    start
  else
    NA
}
```

Test:

```{r}
first_unique(c(1,2,3,4), 4)
first_unique(c(1,1,3,4), 4)
first_unique(c(1,1,3,4,5), 4)
```

Works!

The next function is because I misread the question and return the start rather than end position, but I like `first_unique` so I'm going to keep it as is :)

```{r}
first_marker <- function(x, len)
  first_unique(x, len) + len - 1
```


These all work:

```{r}
first_marker("bvwbjplbgvbhsrlpgdmjqwftvncz" |> str2vec(), 4)
```

```{r}
first_marker("nppdvjthqldpwncqszvftbrmjlhg" |> str2vec(), 4)
```

```{r}
first_marker("nznrnfrfntjfmvfwmzdfjlvtqnbhcprsg" |> str2vec(), 4)
```

So, the answer to Part 1 is:

```{r}
first_marker(vec, 4)
```

Part 2:

```{r}
first_marker("nppdvjthqldpwncqszvftbrmjlhg" |> str2vec(), 14)
```
Works - so Part 2 is:

```{r}
first_marker(vec, 14)
```

We're done!


# Another go

After finishing, I discovered the `zoo::rollapply` function [via Mastodon](https://fosstodon.org/@ashbaldry/109465486319687847), which would have significantly accelerated the solution.

```{r message=FALSE, warning=FALSE}
library(zoo)
```

```{r}
first_marker_2 <- function(x, len) {
  which(rollapply(x, len, all_unique))[1] + len - 1
}
```

Part 1:

```{r}
first_marker_2(vec, 4)
```

Part 2:

```{r}
first_marker_2(vec, 14)
```
