---
title: "Day 5: Supply Stacks"
author: "@Andi@tech.lgbt"
output: 
  html_notebook: 
    code_folding: none
---

```{r}
library(tidyverse)
```

## Sort out the data

```{r}
raw_dat <- read_lines("aoc05.txt")
head(raw_dat, 20)
```

Well, what a mess!

I'll begin by separating the state from the move data.

```{r}
blanks <- which(raw_dat == "")
stopifnot(length(blanks) >= 1)
```

```{r}
state_raw <- raw_dat[1:(blanks[1] - 2)]
state_raw
```

That state representation is going to be tedious to fix. I want it to be something like a list of stacks.

### Moves data

The moves data is easier:

```{r}
moves_raw <- raw_dat[(blanks[1] + 1):length(raw_dat)]
moves_raw |> head(5)
```

```{r}
moves <- moves_raw |>
  str_split_fixed(" ", 6) |>
  as_tibble() |>
  select(V2, V4, V6) |>
  rename(
    n    = V2,
    from = V4,
    to   = V6
  ) |>
  mutate_all(as.integer)
head(moves)
```

### State data

This is dirty.

```{r}
mess_width <- str_length(state_raw[1])
mess_width
```

```{r}
half_fixed <- str_split_fixed(state_raw,
                              pattern = "",
                              n = 35)[, seq(2, mess_width, 4)]
half_fixed
```

```{r}
puzzle_state <- list()
for (i in 1:ncol(half_fixed)) {
  puzzle_state[[i]] <- half_fixed[,i] |> rev()
  puzzle_state[[i]] <- puzzle_state[[i]][puzzle_state[[i]] != " "]
}
puzzle_state
```

## Stack

I'll represent as stack as a vector.

```{r}
empty_stack <- function() {
  c()
}
```

```{r}
is_empty <- function(s) {
  length(s) == 0
}
```

```{r}
is_empty(empty_stack())
```

```{r}
push <- function(s, e) {
  stopifnot(length(e) == 1)
  c(s, e)
}
```

```{r}
test_stack <- push(empty_stack(), "a") |> push("b") |> push("c")
test_stack
```

```{r}
pop <- function(s) {
  stopifnot(length(s) > 0)
  head(s, length(s) - 1)
}
```

```{r}
top <- function(s) {
  stopifnot(length(s) >= 1)
  s[length(s)]
}
```

```{r}
top(test_stack)
```
```{r}
pop(test_stack)
```

```{r}
move <- function(source, target) {
  new_target <- target |> push(top(source))
  new_source <- pop(source)
  list(source = new_source, target = new_target)
}
```


```{r}
move(c("a","b"), c("c","d"))
```

```{r}
move_n <- function(n, source, target) {
  stopifnot(n >= 1)
  
  new_source <- source
  new_target <- target
  
  for (i in 1:n) {
    res_pair   <- move(new_source, new_target)
    new_source <- res_pair$source
    new_target <- res_pair$target
  }
  
  list(source = new_source, target = new_target)  
}
```


```{r}
move_n(3, c("a","b","c"), c("d"))
```

```{r}
test_state <- list(c("a", "c", "d"), c(), "z", c())
test_state
```


```{r}
trans_state <- function(start, n, from_i, to_i) {
  res <- move_n(n, start[[from_i]], start[[to_i]])
  end <- start
  end[[from_i]] <- res$source
  end[[to_i]]   <- res$target
  end
}
```


```{r}
trans_state(test_state, 3, 1, 2)
```


```{r}
trans_states <- function(start, ns, from_is, to_is) {
  stopifnot(length(ns) == length(from_is))
  stopifnot(length(from_is) == length(to_is))
  
  state <- start
  
  for (i in 1:length(ns)) {
    state <- trans_state(state, ns[i], from_is[i], to_is[i])
  }
  
  state
}
```


```{r}
test_state
```


```{r}
trans_states(test_state, c(2,1), c(1,2), c(2,4))
```

# Part 1: glued together

```{r}
the_answer <- trans_states(puzzle_state,
                           moves$n,
                           moves$from,
                           moves$to)
the_answer
```

```{r}
length(the_answer)
```

```{r}
print_tops <- function(ss) {
  res <- rep(NA, length(ss))
  
  for (i in 1:length(ss)) {
    res[i] <- top(ss[[i]])
  }
  
  res
}
```

```{r}
print_tops(the_answer) |> paste(collapse = "")
```
YES!


# Part 2

Okay, I'm going to do something unforgivably hacky -- some surgery on `move_n`.

```{r}
rev_tail <- function(x, n) {
  the_head <- head(x, length(x) - n)
  the_tail <- tail(x, n)
  c(the_head, rev(the_tail))
}
```

```{r}
move_n_9001 <- function(n, source, target) {
  stopifnot(n >= 1)
  
  new_source <- source
  new_target <- target
  
  for (i in 1:n) {
    res_pair   <- move(new_source, new_target)
    new_source <- res_pair$source
    new_target <- res_pair$target
  }
  
  list(source = new_source, target = rev_tail(new_target, n))  
}
```

```{r}
move_n <- move_n_9001
```


```{r}
the_answer_9001 <- trans_states(puzzle_state,
                                moves$n,
                                moves$from,
                                moves$to)
the_answer_9001 |> print_tops() |> paste(collapse = "")
```

Phew - it worked!


