---
title: "Day 8: Treetop Tree House"
author: "@Andi@tech.lgbt"
output: 
  html_notebook: 
    code_folding: none
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
dat <- read.table("aoc08.txt", colClasses = "character")
```


```{r}
dims <- dat[1,] |> str_length()
stopifnot(dims == nrow(dat))
dims
```

```{r}
trees <- str_split_fixed(dat$V1, "", dims)
str(trees)
```

The following function takes a column or a row (it doesn't mind) and tree position, and answers whether the tree at that position is visible to the left/right or top/bottom.

```{r}
is_vis <- function(col, i) {
  col <- as.numeric(col)
  max_left  <- max(col[1:(i - 1)])
  max_right <- max(col[(i+1):length(col)])
  (col[i] > max_left) || (col[i] > max_right)
}
```

`vis_edges` then uses it across all cells. We get the outer cell answers for free, as per the first line below:

```{r}
vis_edges <- function(mat) {
  vis <- nrow(mat) * 2 + (ncol(mat) - 2) * 2

  for (r in 2:(nrow(mat) - 1))
    for (c in 2:(ncol(mat) - 1)) {
      vis_row <- is_vis(mat[r,], c)
      vis_col <- is_vis(mat[,c], r)
      
      if (vis_row || vis_col)
        vis <- vis + 1
    }
  
  vis
}
```

Part 1:

```{r}
vis_edges(trees)
```
It worked (second time -- made an error indexing rows/cols which wasn't picked up on the example).


Part 2:

This function solves it in one direction from a particular tree height.

```{r}
halfcol <- function(col, height) {
  views <- 0
  i <- 1
  blocked <- FALSE
  
  while (i <= length(col) && !blocked) {
    views <- views + 1
    if (col[i] < height)
      i <- i + 1
    else
      blocked <- TRUE
  }

  views
}
```

The following function solves it for a particular row/column:

```{r}
scol <- function(col, i) {
  col <- as.numeric(col)
  
  left_res  <- halfcol(rev(col[1:(i-1)]), col[i])
  right_res <- halfcol(col[(i+1):length(col)], col[i])
  
  left_res * right_res
}
```

Finally, glue it altogether. Edges are zero, hence initialising the matrix result, res, with zeros.

```{r}
scenic_scores <- function(mat) {
  res <- matrix(0,
                ncol = ncol(mat),
                nrow = nrow(mat))

  for (r in 2:(nrow(mat) - 1))
    for (c in 2:(ncol(mat) - 1))
      res[r,c] <- scol(mat[r,], c) *
                  scol(mat[,c], r)
  
  res
}
```


```{r}
scores <- scenic_scores(trees)
```


```{r}
max(scores)
```

Done.



