---
title: "Day 7: No Space Left On Device"
author: "@Andi@tech.lgbt"
output: 
  html_notebook: 
    code_folding: none
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
cmds <- read_lines("aoc07.txt")
str(cmds)
```

```{r}
dat <- tibble(
  cmds = cmds
)
```

```{r}
split_space <- function(s) {
  strsplit(s, split = " ")[[1]]
}
```

I'm starting by getting the absolute paths at each point in the command/output stream:

```{r}
get_abs_path <- function(mess) {
  res <- rep(NA, length(mess))
  where_am_i <- NA
  
  for (i in seq_along(mess)) {
    if (mess[i] |> str_starts("\\$ cd")) {
      dir <- split_space(mess[i])[3]
      if (dir == "/") {
        where_am_i <- c()
      }
      else if (dir == "..") {
        stopifnot(!any(is.na(where_am_i)) &&
                    (length(where_am_i) > 0))
        where_am_i <-
          head(where_am_i,
               length(where_am_i) - 1)
      }
      else {
        stopifnot(!any(is.na(where_am_i)))
        where_am_i <- c(where_am_i, dir)
      }
    }
    res[i] <- paste0("/", paste(where_am_i, collapse = "/"))
  }
  
  res
}
```


```{r}
dat$path <- get_abs_path(dat$cmds)
```

Next, the filesizes - any output beginning with a number.

```{r}
get_filesize <- function(mess) {
  res <- rep(NA, length(mess))

  for (i in seq_along(mess)) {
    if (mess[i] |> str_starts("\\d")) {  
       res[i] <- split_space(mess[i])[1] |> as.numeric()
    }
  }
  
  res
}
```

```{r}
dat$size <- get_filesize(dat$cmds)
```

```{r}
dat$size |> na.omit() |> sum()
```
Now we just do some sums.

```{r}
dat_size <- dat |>
  group_by(path) |>
  summarise(sum_size = sum(size, na.rm = TRUE))
dat_size
```

```{r}
sum(dat_size$sum_size)
```

Totals are easy to find by summing folder sizes for all absolute paths that start with the current path.

```{r}
get_total_sizes <- function(totsdat) {
  res <- rep(NA, nrow(totsdat))

  for (i in seq_along(totsdat$sum_size)) {
    matches <- totsdat |>
      filter(str_starts(path, path[i]))
    res[i] <- sum(matches$sum_size)
  }
  
  res 
}
```

```{r}
dat_size$total_size <- get_total_sizes(dat_size)
```

```{r}
dat_size |>
  select(path, sum_size, total_size)
```

```{r}
most100000 <- dat_size |>
  filter(total_size <= 100000)
most100000 
```


Part 1:

```{r}
most100000$total_size |> sum()
```


Part 2:

```{r}
total_space <- sum(dat_size$sum_size)
total_space
```

"The total disk space available to the filesystem is 70000000. To run the update, you need unused space of at least 30000000."

```{r}
dat_size$left_if_delete <- 70000000 - total_space + dat_size$total_size
```

```{r}
delete_em <- dat_size |>
  filter(left_if_delete >= 30000000) |>
  arrange(left_if_delete)
delete_em
```

So, the answer is:

```{r}
(delete_em |> pull(total_size))[1]
```

