---
title: "Day 11: Monkey in the Middle"
author: "@Andi@tech.lgbt"
output: 
  html_notebook: 
    code_folding: none
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
monkey_mess <- read_lines("aoc11.txt")
str(monkey_mess)
```

```{r}
head(monkey_mess,10)
```

What a mess, `monkey_mess`.

```{r}
less_mess <- tibble(raw = monkey_mess) |>
  filter(raw != "") |>
  mutate(m = sort(rep(0:((n()/6) - 1), n()/(n()/6)))) |> # LOL
  filter(!str_starts(raw, "Monkey")) |>
  mutate(name = rep(c("items", "op", "test", "true", "false"), n()/5),
         rawish = str_split_fixed(raw, ": ", 2)[,2]) |>
  pivot_wider(id_cols = m, names_from = name, values_from = rawish)
less_mess
```

```{r}
lop_off <- function(s, l) {
  str_split_fixed(s, l, 2)[,2]
}
```

```{r}
mm <- less_mess |>
  mutate(op    = lop_off(op, "new = "),
         test  = lop_off(test, "divisible by ") |> as.numeric(),
         true  = lop_off(true, "throw to monkey "),
         false = lop_off(false, "throw to monkey "))
mm
```

That's better.

We can also get the `eval` function out of the way:

```{r}
eval_op <- function(old, str) {
  eval(parse(text = str))
}

eval_op(6, "old * 7")
```
Thankfully R's `eval` and `parse` functions mean I don't need to write a parser.

So, roughly speaking, I want to represent the monkeys' states as a list of lists. I'll write a function that transforms the states for one round and then iterate that as many times as required.


```{r}
ms <- mapply(\(x) strsplit(x, ", "), mm$items) |>
      lapply(as.numeric)
names(ms) <- mm$m
ms
```


This function is horrible. I initially wrote and tested it for one round, and then just wrapped it in a big ol' for-loop for *n*. No regrets.

```{r}
monkey_round <- function(start, n) {
  state   <- start
  inspect <- rep(0, length(ms)) |> as.list()
  names(inspect) <- names(state)

  for (j in 1:n) {
    for (this_monkey in mm$m) {
      cur_row <- mm |> filter(this_monkey == m)
      m_char <- as.character(this_monkey)
      
      for (i in state[[m_char]]) {
        state[[m_char]] <- tail(state[[m_char]],
                                length(state[[m_char]]) - 1)
        i <- eval_op(i, cur_row$op)
        i <- floor(i/3)
        
        if (i %% cur_row$test == 0) {
          state[[cur_row$true]] <- c(state[[cur_row$true]], i)
        }
        else {
          state[[cur_row$false]] <- c(state[[cur_row$false]], i)
        }
        
        inspect[[m_char]] <- inspect[[m_char]] + 1
      }
    }
  }
  
  list(state = state, inspect = inspect)
}
```



## Part 1:

```{r}
res <- monkey_round(ms, 20)
res
```


```{r}
inspections <- sapply(res$inspect, first) |> sort(decreasing = T)
inspections[1] * inspections[2]
```



## Part 2:

The numbers get big, very big. I'd spotted the prime mods and realised that was significant, but needed the explanation [on Reddit](https://www.reddit.com/r/adventofcode/comments/zihouc/comment/izrimjo/?utm_source=share&utm_medium=web2x&context=3).



```{r}
the_mod <- prod(mm$test) # Thanks Reddit!
the_mod
```


```{r}
monkey_round_2 <- function(start, n) {
  state   <- start
  inspect <- rep(0, length(ms)) |> as.list()
  names(inspect) <- names(state)

  for (j in 1:n) {
    for (this_monkey in mm$m) {
      cur_row <- mm |> filter(this_monkey == m)
      m_char <- as.character(this_monkey)
      
      for (i in state[[m_char]]) {
        state[[m_char]] <- tail(state[[m_char]],
                                length(state[[m_char]]) - 1)
        i <- eval_op(i, cur_row$op) %% the_mod
                                    # Thanks Reddit!
        
        if (i %% cur_row$test == 0) {
          state[[cur_row$true]] <- c(state[[cur_row$true]], i)
        }
        else {
          state[[cur_row$false]] <- c(state[[cur_row$false]], i)
        }
        
        inspect[[m_char]] <- inspect[[m_char]] + 1
      }
    }
  }
  
  list(state = state, inspect = inspect)
}
```



```{r}
res2 <- monkey_round_2(ms, 10000)
inspections <- sapply(res2$inspect, first) |> sort(decreasing = T)
inspections[1] * inspections[2]
```



