---
title: "Day 18: Boiling Boulders"
author: "@Andi@tech.lgbt"
output: 
  html_notebook: 
    code_folding: none
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tictoc)
```


```{r}
myCubeColour <- adjustcolor("#ff1199", alpha.f = 0.8)
```


```{r}
dat <- read.table("aoc18.txt", sep = ",") |>
  rename(x = V1, y = V2, z = V3) |>
  mutate(nb = paste(x,y,z,sep = ","))
head(dat)
```


```{r}
neighbours <- function(x, y, z) {
  tibble(
    x = c(x - 1, x + 1,     x,     x,     x,     x),
    y = c(    y,     y, y - 1, y + 1,     y,     y),
    z = c(    z,     z,     z,     z, z - 1, z + 1)
  )
}

neighbours_str <- function(x, y, z) {
  neighbours(x,y,z) |> mutate(str = paste(x,y,z,sep = ",")) |> pull(str)
}
```


```{r}
neighbours_str(0,0,0)
```

```{r}
gen_neighbours <- function(dat) {
  res <- tibble()
  
  for (r in 1:nrow(dat)) {
    res <- rbind(res,
                 tibble(
                   nb = neighbours_str(dat[r, ]$x,
                                       dat[r, ]$y,
                                       dat[r, ]$z),
                   x = dat[r, ]$x,
                   y = dat[r, ]$y,
                   z = dat[r, ]$z
                 ))
  }
  res
}
```




## Part 1:

```{r}
tic()
the_neighbours <- gen_neighbours(dat)
toc()
```

```{r}
head(the_neighbours)
```


The answer is...

```{r}
the_answer <- the_neighbours |> filter(!nb %in% dat$nb)
the_answer |> nrow()
```

3494 is correct.


## Part 2:

Let's have a look...

```{r}
#remotes::install_github('coolbutuseless/isocubes')
library(isocubes)
library(grid)
library(purrr)
```

```{r}
grid.newpage()
isocubesGrob(dat, ysize = 1/40, fill = myCubeColour) |>
  grid.draw() 
```


So, the plan for this one -- another graph, this time undirected. Plop a cuboid over the top of the shape above, remove the shape, leaving space, and then setup a graph that represents all adjacent space cubes. Finally, for each of those cubes, see if it's possible to reach somewhere outside, say the mins of x, y, and z.


The following makes the space:

```{r}
empty_cube <- function(dat) {
  x_ran <- c(min(dat$x) - 1, max(dat$x) + 1)
  y_ran <- c(min(dat$y) - 1, max(dat$y) + 1)
  z_ran <- c(min(dat$z) - 1, max(dat$z) + 1)
  
  expand.grid(x = x_ran[1]:x_ran[2],
              y = y_ran[1]:y_ran[2],
              z = z_ran[1]:z_ran[2]) |>
    mutate(nb = paste(x, y, z, sep = ",")) |>
    filter(!nb %in% dat$nb)
}
```

```{r}
frame_around <- empty_cube(dat)
```


Now the graph action.

```{r message=FALSE, warning=FALSE}
library(igraph)
```



```{r}
buildgraph <- function(dat) {
  res <- data.frame()
  
  for (r in 1:nrow(dat)) {
    neighbours <- neighbours_str(dat[r,]$x, dat[r,]$y, dat[r,]$z)
    
    there_str <- dat |>
      filter(nb %in% neighbours) |>
      pull(nb)
    
    here_str  <-
      rep(paste(dat[r,]$x, dat[r,]$y, dat[r,]$z, sep = ","),
          length(there_str))
    
    res <-
      rbind(res, data.frame(here = here_str, there = there_str))
  }
  
  res |> unique() |> graph.data.frame(directed = FALSE)
}
```


```{r}
tic()
space_graph <- buildgraph(frame_around)
toc()
```



```{r}
all_outside <- function(the_graph, the_space_dat) {
  res <- the_space_dat
  res$outside <- NA
  vs <- attr(V(space_graph), "names")
  
  epitome_of_out <- c(min(the_space_dat$x),
                      min(the_space_dat$y),
                      min(the_space_dat$z)) |> paste(collapse = ",")
  
  for (r in 1:nrow(res)) {
    there_str <- res[r,]$nb
    if (!there_str %in% vs)
      res$outside[r] <- FALSE
    else
      res$outside[r] <-
        !is.infinite(distances(the_graph, epitome_of_out, there_str, mode = "out"))
  }
  
  res
}
```


```{r}
tic()
inside <- all_outside(space_graph, frame_around) |>
  filter(!outside)
toc()
```


```{r}
the_answer <- the_neighbours |> filter(!nb %in% dat$nb &
                                       !nb %in% inside$nb)
the_answer |> nrow()
```

YES.
