---
title: "Day 15: Beacon Exclusion Zone - part 1 works"
author: "@Andi@tech.lgbt"
output: 
  html_notebook: 
    code_folding: none
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
read_sensor_dat <- function(filename) {
  tibble(str = read_lines(filename)) |>
    extract(
      str,
      into = c("Sx", "Sy", "Bx", "By"),
      regex = "Sensor at x=(-?[0-9]+), y=(-?[0-9]+): closest beacon is at x=(-?[0-9]+), y=(-?[0-9]+)"
    ) |>
    mutate_all(as.numeric) |>
    mutate(
      d = manhat(Sx, Bx, Sy, By)
    )
}
```

```{r}
manhat <- function(x1,x2,y1,y2) {
  abs(x1 - x2) + abs(y1 - y2)
}
```

```{r}
testdat <- read_sensor_dat("aoc15test.txt")
testdat 
```

```{r}
scan_row <- function(sensors, here_y, from_x, to_x) {
  gaps   <- 0
  
  for (here_x in from_x:to_x) {
    sensors_here <- sensors |>
      mutate(
        d_here = manhat(here_x, Sx, here_y, Sy),
        closer = d_here <= d,
        beacon_here = here_x == Bx & here_y == By
      )
    
    if (any(sensors_here$closer) && !any(sensors_here$beacon_here))
    {
      gaps <- gaps + 1
    }
  }
  
  gaps
}
```

```{r}
solve_it <- function(dat, y) {
  scan_row(
    dat,
    here_y = y,
    from_x = min(dat$Sx - dat$d),
    to_x   = max(dat$Sx + dat$d)
  )
}
```

The right answer is 26:

```{r}
solve_it(read_sensor_dat("aoc15test.txt"), y = 10)
```
Worked!


## Part 1:

It's a much bigger search space but I reckon it'll work...

```{r eval=FALSE}
solve_it(read_sensor_dat("aoc15.txt"), y = 2000000)
```
5144286

Yes! Eventually (~4 hours).


## Part 2:

Try brute forcing again for the small one.

```{r}
scan_rect <- function(sensors, from_x, to_x, from_y, to_y) {
  res <- NA
  
  for (here_y in from_y:to_y)
    for (here_x in from_x:to_x) {
      sensors_here <- sensors |>
        mutate(
          d_here = manhat(here_x, Sx, here_y, Sy),
          closer = d_here <= d,
          beacon_here = here_x == Bx & here_y == By
        )
      
      if (!any(sensors_here$closer) &&
          !any(sensors_here$beacon_here))
      {
        res <- c(here_x, here_y)
        break
      }
    }
  
  res
}
```


It's just 400 cells, so this will work.

```{r}
foundit <- scan_rect(testdat, 0, 20, 0, 20)
foundit
```

```{r}
foundit[1] * 4000000 + foundit[2]
```

This is correct, so at least I'm confident I get the problem spec.

The full problem has an absurdly large search space so this isn't going to work...

```{r}
4000000^2
```
Let's not.


## Some pictures

```{r}
realdat <- read_sensor_dat("aoc15.txt")
```



```{r}
drawpolys <-
  function(dat,
           xlim,
           ylim,
           col = "grey",
           border = "black",
           outside = "black",
           xsearchlim,
           ysearchlim) {
    plot(
      1,
      xlim = xlim,
      ylim = ylim,
      xlab = "x",
      ylab = "y",
      type = "n"
    )
    polygon(
      x = c(xsearchlim[1], xsearchlim[2], xsearchlim[2], xsearchlim[1]),
      y = c(ysearchlim[1], ysearchlim[1], ysearchlim[2], ysearchlim[2]),
      border = outside,
      col = "yellow"
    )
    
    for (r in 1:nrow(dat)) {
      x_c   <- dat[r,]$Sx
      y_c   <- dat[r,]$Sy
      d_h   <- dat[r,]$d
      
      polygon(
        x = c(x_c, x_c - d_h, x_c, x_c + d_h),
        y = c(y_c + d_h, y_c, y_c - d_h, y_c),
        col = col,
        border = border
      )
    }
    polygon(
      x = c(xsearchlim[1], xsearchlim[2], xsearchlim[2], xsearchlim[1]),
      y = c(ysearchlim[1], ysearchlim[1], ysearchlim[2], ysearchlim[2]),
      border = outside
    )
    
  }
```




```{r fig.height=4, fig.width=4}
drawpolys(
  realdat,
  c(0, 4000000),
  c(0, 4000000),
  col = adjustcolor("grey",alpha.f=.5),
  border = NA,
  outside = "cyan",
  xsearchlim = c(0,4e6),
  ysearchlim = c(0,4e6)
)
```



```{r fig.height=4, fig.width=4}
drawpolys(
  testdat,
  c(-10, 30),
  c(-10, 30),
  col = adjustcolor("grey", alpha.f = .5),
  border = "black",
  outside = "cyan",
  xsearchlim = c(0, 20),
  ysearchlim = c(0, 20)
)
points(14, 11, pch = 19, col = "red")
```



To be continued. This looks like an interesting problem to study.





