---
title: "Day 13: Distress Signal"
author: "@Andi@tech.lgbt"
output: 
  html_notebook: 
    code_folding: none
---


```{r message=FALSE, warning=FALSE}
library(tidyverse)
```


```{r}
read_packets <- function(filename) {
  dat <- tibble(str = read_lines(filename)) |>
    filter(str != "") |>
    mutate(pair = rep(c("L", "R"), n() / 2),
           i    = rep(1:(n() / 2), 2) |> sort()) |>
    pivot_wider(values_from = str, names_from = pair)
  dat
}
```


```{r}
ps <- read_packets("aoc13test.txt")
ps
```

Yep, I am going to use `gsub`, `parse`, and `eval` to turn those list strings into R lists :)

```{r}
listify <- function(str) {
  l_fixed <- gsub("\\[", "list(", str)
  
  gsub("\\]", ")", l_fixed) |>
    parse(text = _) |>
    eval()
}
```

```{r}
listify("[1,2,3,[4,5], 6]")
```

To help make debugging less painful:

```{r}
print_list <- function(l) {
  if (!is.list(l))
    cat(l)
  else {
    cat("[")
    cat("")
    i <- 1
    for (e in l) {
      print_list(e)
      if (i == length(l))
        break;
      cat(",")
      i <- i + 1
    }
    cat("]")
  }
}
```

```{r}
print_list(list(list(1,2),list()))
```

I got the following function to work for the test input, but not the actual puzzle.

Thanks [@jrh-dev](https://github.com/jrh-dev) for the [working logic](https://github.com/jrh-dev/Advent-of-Code-2022/blob/main/R/d13.r), slotted into `packet_lists_ordered`...

```{r}
packet_lists_ordered <- function(l, r, verbose = FALSE) {
  if (verbose) {
    print_list(l)
    cat("\n")
    print_list(r)
    cat("\n\n")
  }
  
  if (length(l) == 0 && length(r) == 0)
    return(NA)
  else if (length(l) == 0)
    return(TRUE)
  else if (length(r) == 0)
    return(FALSE)
  
  if (is.numeric(l[[1]]) && is.numeric(r[[1]])) {
    if (l[[1]] < r[[1]])
      return(TRUE)
    else if (l[[1]] > r[[1]])
      return(FALSE)
    else {
      l[[1]] = r[[1]] = NULL
      return(packet_lists_ordered(l, r))
    }
  }
  
  if (is.list(l[[1]]) && is.list(r[[1]])) {
    tmp = packet_lists_ordered(l[[1]], r[[1]])
    if (is.na(tmp)) {
      l[[1]] = r[[1]] = NULL
      return(packet_lists_ordered(l, r))
    } else {
      return(tmp)
    }
  }
  
  if (is.numeric(l[[1]]) & is.list(r[[1]])) {
    l[[1]] <- list(l[[1]])
    return(packet_lists_ordered(l, r))
  }
  
  if (is.list(l[[1]]) & is.numeric(r[[1]])) {
    r[[1]] <- list(r[[1]])
    return(packet_lists_ordered(l, r))
  }
}
```


```{r}
packet_strs_ordered <- function(str1, str2) {
  left  <- listify(str1)
  right <- listify(str2)
  
  packet_lists_ordered(left, right)
}
```


```{r}
ps$ordered <- mapply(packet_strs_ordered, ps$L, ps$R)
ps |> filter(ordered) |> pull(i)
```

```{r}
stopifnot(ps |> filter(ordered) |> pull(i) |> sum() == 13)
```

If we got this far then the code worked for the example.

Onto the actual input.

## Part 1:

```{r}
ps_actual <- read_packets("aoc13.txt")
```

```{r}
ps_actual$ordered <-
  mapply(packet_strs_ordered, ps_actual$L, ps_actual$R)
```

```{r}
ps_actual |> filter(ordered) |> pull(i) |> sum()
```
Yep -- thanks to [@jrh-dev](https://github.com/jrh-dev).



## Part 2:

Setup the input to be sorted.

```{r}
sort_em <- ps_actual |> pivot_longer(cols = c(L,R)) |> pull(value)
sort_em <- c(sort_em, "[[2]]", "[[6]]")
```

It's a small input, so this will do...

```{r}
sink_sort <- function(pstrs) {
  res <- pstrs
  for (i in 1:(length(res) - 1))
    for (j in (i + 1):length(res))
      if (!packet_strs_ordered(res[i], res[j])) {
        tmp <- res[i]
        res[i] <- res[j]
        res[j] <- tmp
      }

  res
}
```


```{r}
sorted <- sink_sort(sort_em)
```

```{r}
which(sorted == "[[2]]") * which(sorted == "[[6]]")
```

Hurrah.
