---
title: "Day 12: Hill Climbing Algorithm"
author: "@Andi@tech.lgbt"
output: 
  html_notebook: 
    code_folding: none
---


```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

Read in the map:

```{r}
load_map <- function(filename) {
  rawdat <- read_lines(filename)
  str_split_fixed(rawdat, "", rawdat[1] |> str_length())
}
```

```{r}
testdat <- load_map("aoc12test.txt")
testdat
```



## Chancing my arm with a brute force recursive approach

Guess what happened...?


```{r}
climbable <- Vectorize(function(here, there) {
  if (is.na(here) || is.na(there))
    FALSE
  else {
    here_n  <- which(letters == here)
    there_n <- which(letters == there)
    there_n - here_n <= 1
  }
})
```


```{r}
neighbours <- function(mat, where) {
  w.r  <- where[1]
  w.c  <- where[2]
  here <- mat[w.r, w.c]
  
  res <- data.frame(r = c(w.r,     w.r,     w.r - 1, w.r + 1),
                    c = c(w.c - 1, w.c + 1, w.c,     w.c)) |>
    filter(r >= 1 &
           c >= 1 &
           r <= nrow(mat) &
           c <= ncol(mat)) |>
    filter(!is.na(mat[cbind(r, c)])) |>
    mutate(there = mat[cbind(r, c)])
  
  if (nrow(res) > 0)
    res |> filter(climbable(here, there))
  else
    res
}
```


```{r}
find_path_worker <- function(start, end, mat) {
  cur_mat <- mat
  cur_mat[start] <- NA
  res <- NA
  
  if (all(start == end)) {
    res <- 0
  }
  else {
    venture <- neighbours(mat, start)
    
    if (nrow(venture) == 0) {
      res <- NA
    }
    else {
      venture$dist <- mapply(\(r,c) find_path_worker(cbind(r,c),
                                                     end,
                                                     cur_mat),
                             venture$r,
                             venture$c) 
      venture <- na.omit(venture)
      if (nrow(venture) == 0)
        res <- NA
      else
        res <- 1 + min(venture$dist)
    }
  }
  
  res
}
```


```{r}
find_path <- function(mat) {
  S_is <- which(mat == "S", arr.ind = TRUE)
  E_is <- which(mat == "E", arr.ind = TRUE)
  
  mat[S_is] <- "a"
  mat[E_is] <- "z"
  
  find_path_worker(S_is, E_is, mat)
}
```


It actually works for the test example, after a minute or two!

```{r}
find_path(testdat)
```

LOL. Not a hope on the full problem.


## Time for a graph algorithm

I hadn't used graph algorithms in R before. A Google for "graph shortest distance r" led to {igraph}. Fingers crossed.


```{r message=FALSE, warning=FALSE}
library(igraph)
```


I'm going to reuse `neighbours` to build the DAG. For node names I'll use the coordinates as a comma-separated string.


```{r}
prep_dat <- function(mat) {
  S_is <- which(mat == "S", arr.ind = TRUE)
  E_is <- which(mat == "E", arr.ind = TRUE)
  
  mat[S_is] <- "a"
  mat[E_is] <- "z"
  
  list(input = mat, S = S_is, E = E_is)
}
```




```{r}
build_edges <- function(mat) {
  res <- data.frame()
  
  for (r_here in 1:nrow(mat)) {
    for (c_here in 1:ncol(mat)) {
      to_dat <- neighbours(mat, c(r_here,c_here)) |>
        mutate(from = paste(r_here, c_here, sep = ","),
               to   = paste(r, c, sep = ",")) |>
        select(from, to)
      res <- rbind(res, to_dat)
    }
  }
  res
}
```



```{r}
build_edges(prep_dat(testdat)$input)
```



```{r}
edges_to_dag <- function(edges) {
  graph.data.frame(d = edges, directed = TRUE)
}
```



```{r}
test_dag <- build_edges(prep_dat(testdat)$input) |>
            edges_to_dag()
```



```{r}
plot(test_dag)
```


```{r}
start_edge <- function(mat) {
  which(mat == "S", arr.ind = TRUE) |> paste(collapse = ",")
}
end_edge <- function(mat) {
  which(mat == "E", arr.ind = TRUE) |> paste(collapse = ",")
}
```


```{r}
start_edge(testdat)
end_edge(testdat)
```



```{r}
test_paths <- shortest_paths(
  graph  = test_dag,
  from   = start_edge(testdat),
  to     = end_edge(testdat),
  output = "epath"
)
```


```{r}
test_paths$epath[[1]] |>length()
```

Blimey that was fast. Let's try the actual puzzle.



```{r}
realdat <- load_map("aoc12.txt")
str(realdat)
```

```{r}
real_dag <- build_edges(prep_dat(realdat)$input) |>
            edges_to_dag()
```



```{r}
real_paths <- shortest_paths(
  graph  = real_dag,
  from   = start_edge(realdat),
  to     = end_edge(realdat),
  output = "epath"
)
```


### Part 1:

```{r}
real_paths$epath[[1]] |>length()
```

What a relief -- that worked.



### Part 2:

I think I can reuse the DAG.




```{r}
the_starts <- which(prep_dat(realdat)$input == "a", arr.ind = TRUE)
head(the_starts)
```


```{r}
the_end <- which(realdat == "E", arr.ind = TRUE)
the_end
```



```{r}
path_len <- function(dag, start, end) {
  the_paths <- shortest_paths(
    graph  = dag,
    from   = paste(start, collapse = ","),
    to     = paste(end,   collapse = ","),
    output = "epath"
  )
  the_paths$epath[[1]] |>length()  
}
```

```{r}
dists <- rep(NA, nrow(the_starts))
```

Warnings suppressed below... it's all good.

```{r warning=FALSE}
for (r in 1:nrow(the_starts)) {
  dists[r] <- path_len(real_dag, the_starts[r,], the_end)
}
```



```{r}
dists[which(dists  > 0)] |> min()
```

Hurrah!




